|

### 2020-03-06 analysis/paper/response_letter.Rmd

This is a clearly written and organized manuscript that presents an application of novel, open-source tools to disseminate stock assessment inputs and results. The interactive visualization tools were developed by others but modified by the authors to suit their needs; the original developers are credited appropriately. The authors make a strong case for using these tools for visualization of stock assessment results, and in that way, this paper is fitting for a perspective since it promotes a new paradigm for dissemination of stock assessments. There are no new research results, nor is there anything controversial in this piece. An important and laudable message of the paper is that these tools can help the transition to "a culture of open science and reproducible research."

*We thank the Reviewer for the kind comments.*

The manuscript repeatedly mentions the usefulness of these tools for stakeholders, both for data exploration and data modeling. In the data exploration figures (Figures 1 through 3), the interfaces would be more digestible to stakeholders if they included an annotation of what the user is viewing. For example, the Figure 1 view could include a note specifying that the error bars represent 95% confidence intervals, and species names should be used rather than, or in addition to, codes. Figure 2 could be improved by including units for the length bins, and Figure 3 should have a legend for the kernals. Each could also have a text field that gives an overview of the what the figure is showing and how a user can interact with the figure.

*TODO?: Greg & Bob*

More broadly, the NCAM Explorer (Supplement 2) would also be more functional with increased annotation. The background tab provides important context but should be expanded beyond the "terse point-form" to serve as a user guide for each tab. These improvements apply to the tools themselves rather than to the manuscript.

*We agree with these comments and have added text to the Background tab and added references for background.  See "stand alone" comment above.  We have tried to strike a balance between addressing the spirit of the Reviewers comments and unnecessarily cluttering the dashboard.*

The one area where I think the manuscript should be expanded is in the "Towards open stock-assessment" section. The main reason given for increasing the transparency of stock assessment through tools such as these is to improve peer review. However, as mentioned at several other points in the manuscript, these tools are also valuable for increasing stakeholder engagement in and understanding of stock assessment. That function warrants equal attention to peer review in this section and could present ideas for how to introduce these tools to stakeholders and how to help them sift through the large volume of data to find what is most relevant to them.

*TODO?: Keith*

Lastly, Supplement 1 (getting_started.html) isn’t referenced anywhere in the paper. It can be removed from the supplementary files.

*We have now added a reference to Supplement 1 to the introduction and conclusion.*

Reviewer: 3

Comments to the Author

This manuscript illustrates a novel approach to exploring and communicating data and model results related to fisheries stock assessments. Communication methods for complex patterns in data or stock assessment results is not a frequent topic covered in fisheries science journals but it is an area where new ideas are rapidly developing, much further progress is needed, and greater prominence in the literature is deserved. I highly recommend publishing this manuscript.

*We thank the Reviewer for the kind comments.*

The availability of the NCAM Explorer file NCAM_dashboard.html (referenced in the manuscript as Supplement 2) is vital for the understanding of the interactive nature of the tools developed. It would be useful to reference the other supplemental file getting_started.html in the text and include as an additional supplement an example of the RStrap Explorer as well in addition to the static figures in the text.

*We agree with the Reviewers comments and we have included a reference to Supplement 1 in the introduction and conclusion. RStrap Explorer is harder to include as a supplement as it is Shiny based and must be hosted on a server. At the moment, the use of this tool has been limited to local computers; we have yet to host the tool on an online server.*

I raise a few minor specific issues below, but the area where I would suggest the greatest change is adding to the introduction and conclusion further discussion of a few larger questions that this new approach raises. First, the visualization and communication needs are very broad within the stock assessment community, but the example tools presented are specific to analyses used in the region where the authors work: the RStrap analysis of survey data and the NCAM assessment model. I would have like to see some discussion about the trade-offs associated with developing generic dashboards that could be applied to other survey or assessment tools vs. developing separate dashboard applications for each database or stock assessment model. Comparing results from multiple stock assessment models which might be applied to the same stock is another challenge that might be discussed.

*Please see previous comments to the AE.*

Second, I would have liked more discussion of version control and archiving. The supplemental file NCAM_dashboard.html includes a comparison of two models "NCAM 2018" and "NCAM 2019" but in the assessment process there is typically a proliferation of numerous alternative models being considered in any given year. Exploration of these alternative models would presumably lead to a proliferation of these html dashboard files. Keeping track of which files show the results from which models is a challenge not addressed in the manuscript (and not obvious in the supplemental file). Furthermore, keeping track of the software version for the dashboard software (and R packages on which it depends) and format of the database or assessment model and maintaining compatibility between the two is a challenge not discussed here.

*This is an excellent point, however, we do discuss this issue as this approach does not offer a novel solution to file organization. The fact is that there is little difference between the file management with this approach versus a more standard approach. Through pre-dashboard days, different model runs were nested within independent folders and within these folders were figure and table folders. The only difference with the dashboard approach is that key figures and tables are held in the html file (i.e. there is no longer figure and table folders). A convention we apply to help keep track of things is to use a consistent name for the folder, the RData file containing all the inputs and outputs, and the html file. Though an interesting and important challenge, we did not cover this as file management and project conventions is a potential tangent that may add little to the paper as potential users are likely going to apply their own personal preferences.*

*Regarding software version, we now note in the conclusion that these files can be traceable and accessible via GitHub.*

Third, the link between this tool and more traditional static reports could be also addressed. The manuscript presents a compelling argument in favor of interactive dashboards as a tool for the rich volume of information needing to be explored in fisheries models, but it is not suggested that static reports will go away. It is noted that "The user also has the ability to hover over the points to compare estimates and export the figure as a png file" (line 110) but this would not be an efficient way to produce figures for a static report. Further discussion of how the same tools used for the dashboard view could used to make the development of static reports more efficient may be worth including in the discussion.

*We agree, it would not be very efficient to manually export individual plots. We have removed this note and have expanded our introductory text to clarify how synergies can be created between static and interactive documents.*

Specific issues: Line 29: "For well-monitored stocks, the challenge has shifted from having sufficient data and information for providing sound advice on stock status, to presenting large quantities of data and output from increasingly complex statistical models in a meaningful way." I think this is a mischaracterization of the state of the science. Perhaps Northern Cod is an exception, but I believe the challenge of having adequate information to provide sound advice remains as present as ever, even for relatively well-monitored stocks, while the second of meaningfully presenting the resulting information has grown beside it.

*Fair point. We have rephrased the sentence to read: "While these advances have surely improved the advice provided at stock assessments, it has become more difficult to communicate the basis of this advice in a meaningful way."*

Line 81: the reference to a 1981 report for information on the "locally developed R package called RStrap" seems inadequate. Perhaps "locally developed" means only locally available and thus no citation for the R package is available (in contrast to the other packages cited in the manuscript).

*This is correct.  We now simply state that it is a local package.*

Line 121: "data base" should be one word for consistency with earlier occurrence on the same line.

*We have made the suggested change.*

Lines 262: The statement "Although these tools do not reveal the entire data pipeline" may be contradicted by the ICES Transparent Assessment Framework (https://taf.ices.dk/app/about) which archives the full process of data processing to running assessment models in a reproducible way and should be added to Appendix A.

*We thank the Reviewer for drawing our attention to the TAF. We have now re-structured our "Towards open stock assessment" section to highlight TAF and several other open science initiatives in fisheries research.*

Figure 2: I realize this figure is only for illustrative purposes, but it might be good to add a note about the strange pattern in the length compositions of American plaice in the 1980s. The spiky pattern with peaks in the same bins each year seems like an artifact of rounding error or conversion from some other type of length measurement.

*TODO?: Bob*

Again, I think this manuscript is a valuable contribution to the field, is of broad interest, and I encourage it to be published with or without the refinements proposed above.

*We thank the Reviewer for the kind comments.*
